---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ffsimulator

<!-- badges: start -->
[![CRAN status](https://img.shields.io/cran/v/ffsimulator?style=flat-square&logo=R&label=CRAN)](https://CRAN.R-project.org/package=ffsimulator) [![Dev status](https://img.shields.io/github/r-package/v/dynastyprocess/ffsimulator/main?label=dev&style=flat-square&logo=github)](https://ffsimulator.dynastyprocess.com/dev/) [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg?style=flat-square)](https://lifecycle.r-lib.org/articles/stages.html) [![R build status](https://img.shields.io/github/workflow/status/dynastyprocess/ffsimulator/R-CMD-check?label=R%20check&style=flat-square&logo=github)](https://github.com/DynastyProcess/ffsimulator/actions) [![nflverse discord](https://img.shields.io/discord/591914197219016707.svg?color=5865F2&label=nflverse%20discord&logo=discord&logoColor=5865F2&style=flat-square)](https://discord.com/invite/5Er2FBnnQa)

<!-- badges: end -->

The goal of ffsimulator is to simulate fantasy football seasons. This helps analyze distributions of:

- expected season finish
- player contribution to wins
- etc?

by simulating 100 fantasy seasons with historical weekly finishes + fantasypros current redraft rankings to determine overall/relative team strengths. 

## Installation

Install the development version from GitHub with:

```{r eval = FALSE}
# install.packages("remotes")

remotes::install_github("dynastyprocess/ffsimulator", ref = "dev")
```

## Roadmap

- Connect via ffsimulator
- Download scoring history
- Calculate season ranks and the population of player scores for that rank
- Download latest fantasypros rankings
- Connect rosters to fantasypros rankings and player score outcomes
- randomly select n games for each player
- calculate optimal lineups (learn lpSolve?)
- calculate started lineups
- convert week scores into fantasy season schedule (via tony's ffsched package)
- calculate win/loss (H2H  + allplay) + total season poitns for/potential points
- figure out plots and viz


```{r eval = TRUE}

# library(ffsimulator)
library(ffscrapr)
library(dplyr)
library(tidyr)
library(purrr)
library(arrow)

## CONNECT TO LEAGUE

conn <- mfl_connect(2021, 54040)

## DOWNLOAD SCORING HISTORY

scoring_history <- ff_scoringhistory(conn,2006:2020)

## CALCULATES SEASON RANKS AND THE POPULATION OF PLAYER SCORES FOR THAT RANK

pos_rank <- scoring_history %>% 
  group_by(season, gsis_id, mfl_id, player_name, pos, team) %>% 
  summarise(
    games = n(),
    ppg = mean(points, na.rm = TRUE)
  ) %>% 
  ungroup() %>% 
  filter(
    games > 4
  ) %>% 
  group_by(season, pos) %>% 
  mutate(rank = rank(desc(ppg),ties.method = "random")) %>% 
  ungroup() %>% 
  arrange(season, pos, rank) %>% 
  left_join(
    scoring_history %>% select(season, gsis_id, week, points),
    by = c("season", "gsis_id")
  ) %>% 
  group_by(pos, rank) %>% 
  summarise(
    all_points = list(points)
  ) %>% 
  ungroup() %>% 
  mutate(one_season = map(all_points, sample, size = 1000, replace = TRUE))

# fp_parquet <- tempfile()
# download.file("https://github.com/dynastyprocess/data/raw/master/files/db_fpecr.parquet",fp_parquet)

## DOWNLOADS LATEST FANTASYPROS RANKINGS

fantasypros <- arrow::read_parquet("data-raw/db_fpecr(1).parquet") %>% 
  filter(ecr_type == "ro", scrape_date == max(scrape_date), pos %in% c("QB","RB","WR","TE")) %>% 
  select(player, fantasypros_id = id, pos, team, ecr, sd) %>% 
  group_by(pos) %>% 
  mutate(rank = rank(ecr, ties.method = "random")) %>% 
  ungroup()

## CONNECTS ROSTERS TO FANTASYPROS + PLAYER SCORES 

rosters <- ff_rosters(conn) %>% 
  select(franchise_id, franchise_name, mfl_id = player_id, player_name, pos, team, age) %>% 
  left_join(dp_playerids() %>% select(fantasypros_id, mfl_id),
            by = "mfl_id") %>% 
  left_join(fantasypros %>% select(fantasypros_id, rank),
            by = "fantasypros_id") %>% 
  left_join(pos_rank %>% select(pos,rank,all_points),
            by = c("pos","rank")) %>% 
  mutate(
    season_games = map_if(all_points,!is.na(rank), sample, size = 17, replace = TRUE)
  )

## CALCULATE OPTIMAL LINEUPS

# two paths: a) develop heuristics b) use lpSolve
# ideally solver does the optimisation

starter_positions <- ff_starter_positions(conn)

## LOGIC FOR CALCULATING STARTED LINEUPS
# a) Always choose highest ranked player by fantasypros
# b) Assume random percentage between 70-85% of optimal points

## APPLY WEEK SCORES INTO FANTASY SEASON SCHEDULES
# (use tony's ffsched package)

## CALCULATE WIN/LOSS (H2H, ALL-PLAY) + TOTAL SEASON POINTS FOR/ POTENTIAL POINTS

## figure out plots and visualizations later? 

```

```{r}
redraft_rankings <- read.csv("data-raw/fp_redraft_20162020.csv")

ppg_data <- scoring_history %>% 
  group_by(season, gsis_id, mfl_id, player_name, pos, team) %>% 
  summarise(
    games = n(),
    ppg = mean(points, na.rm = TRUE)
  ) %>% 
  ungroup()

preseason_adp_outcomes <- redraft_rankings %>% 
  filter(!pos %in% c("K","DST")) %>% 
  group_by(season, pos) %>% 
  mutate(pos_rank = rank(rank_ave, ties.method = "random")) %>% 
  ungroup() %>% 
  left_join(
    scoring_history %>% select(sportradar_id, season, week, points) %>% filter(week <=17),
    by = c("season","sportsdata_id"="sportradar_id")
  ) %>% 
  group_by(pos, pos_rank) %>% 
  summarise(
    all_outcomes = list(points),
    n = n()
  ) %>% 
  ungroup() %>% 
  group_by(pos) %>% 
  mutate(
    missing = (5*17) - n,
    missing_na = map(missing, ~rep_len(x = 0, length.out = .x)),
    new_all_outcomes = map2(all_outcomes, missing_na, ~c(...)),
    new_n = map(new_all_outcomes, length),
    extra_outcomes = ifelse(pos_rank == 1, new_all_outcomes, numeric()),
    wide_bins = pmap(list(new_all_outcomes, lead(new_all_outcomes), lag(new_all_outcomes), extra_outcomes),~c(...)),
    wide_n = map(wide_bins, length)
  ) %>% 
  ungroup()

```


```{r}
library(slider)
library(ggridges)
library(hrbrthemes)

preseason_adp_outcomes %>% 
  filter(pos == "RB", pos_rank <= 24) %>% 
  mutate(pos_rank = as.factor(pos_rank)) %>% 
  unnest(wide_bins) %>% 
  mutate(wide_bins = replace_na(wide_bins, 0)) %>%
  ggplot(aes(x = wide_bins, y = pos_rank, fill = pos_rank)) + 
  # geom_density_ridges(colour = "white",quantile_lines = TRUE) + 
  geom_density_ridges(colour = "white",stat = "binline") + 
  theme_modern_rc()

preseason_adp_outcomes %>% 
  mutate(
    sample_weeks = map(wide_bins, ~sample(.x,size = 17,replace = TRUE))
  ) %>% 
  filter(pos == "WR", pos_rank <= 24) %>% 
  mutate(pos_rank = as.factor(pos_rank)) %>% 
  unnest(sample_weeks) %>% 
  ggplot(aes(x = sample_weeks, y = pos_rank, fill = pos_rank)) + 
  # geom_density_ridges(colour = "white", quantile_lines = TRUE) +
  geom_density_ridges(colour = "white",stat = "binline") +
  theme_modern_rc()

outcomes_rb_ppg <- pos_rank %>% 
  select(-one_season) %>% 
  mutate(n = map_dbl(all_points, length)) %>% 
  filter(pos == "RB", rank <= 12)

outcomes_rb <- outcomes_rb_adp %>% 
  rename(all_outcomes = adp_outcomes)

```

